\begin{question}
	\normalfont We say that a matrix $A$ is \emph{similar} to a matrix $B$ if there exists a matrix $Q$ such that 
	\[
	A = Q^{-1}BQ.
	\]
	\begin{enumerate}[(i)]
	\item Prove that if matrices $A$ and $B$ are similar, both $A$ and $B$ must be square matrices, and of the same size. 
	\vspace{.1cm}	
\item	Prove that if $A$ and $B$ are similar matrices, then det($A$) = det($B$). 
	\end{enumerate}	
	\end{question}
\begin{proof}
	\renewcommand{\qedsymbol}{$\blacksquare$}
	\begin{enumerate}[(i)]
		\item Suppose $A$ and $B$ are similar.
		Then, there exists an invertible matrix $Q$ such that $A=Q^{-1}BQ$.
		Since $Q$ is invertible, $Q$ and $Q^{-1}$ must be two $n\times n$ square matrices for some $n\in\N$.
        
		Suppose $B$ is an $a\times b$ matrix.
        Since $A=Q^{-1}BQ$, matrix multiplication must be defined for $Q^{-1}BQ$.
        But since matrix multiplication is associative, $A=(Q^{-1}B)Q$ or $A=Q^{-1}(BQ)$.
        It follows that $Q^{-1}B$ or $BQ$ must be defined.

        If $Q^{-1}B$ is defined, since $Q^{-1}$ is an $n\times n$ matrix and $B$ is an $a\times b$ matrix, $a=n$ and $Q^{-1}B$ is an $n\times b$ matrix.
        But then, $A=(Q^{-1}B)Q$, or $(Q^{-1}B)Q$ must be defined.
        As $Q^{-1}B$ is an $n\times b$ matrix and $Q$ is an $n\times n$ matrix, $b=n$ and $A=(Q^{-1}B)Q$ is an $n\times n$ matrix.
        Hence, $a=b=n$.

        If $BQ$ is defined, since $Q$ is an $n\times n$ matrix and $B$ is an $a\times b$ matrix, $b=n$ and $BQ$ is an $a\times n$ matrix.
        As $BQ$ is an $a\times n$ matrix and $Q^{-1}$ is an $n\times n$ matrix, $a=n$ and $A=Q^{-1}(BQ)$ is an $n\times n$ matrix.
        Hence, $a=b=n$.

        Therefore, $A$ and $B$ are both $n\times n$ square matrices.\qed
        \item From Theorem 3.5 Section 3.3 LADW, for any $n\times n$ matrices $X,Y$,
        \[
            \begin{aligned}
                \det(XY)=\det X\det Y
            \end{aligned}
        \]
        For some $n\times n$ matrix $Q$, since $Q^{-1}Q=I$, where $I$ is the $n\times n$ identity matrix, $\det I=\det(QQ^{-1})=\det Q\det Q^{-1}$, or $\det Q\det Q^{-1}=1$.
        Hence, 
        \[
            \begin{aligned}
                \det A=\det(Q^{-1}BQ)=\det Q^{-1}\det B\det Q=(\det Q^{-1}\det Q)\det B=\det B
            \end{aligned}
        \]
        Therefore, $\det A=\det B$.\qed
     	\end{enumerate}
	\renewcommand{\qedsymbol}{}
\end{proof}